As-you-go instead of after-the-fact: A network approach to scholarly communication and evaluation

### 2018-03-30 docs/manuscript.Rmd
Finalizing manuscript.

---
title: "As-you-go" instead of "after-the-fact": A network approach to scholarly communication and evaluation" 
bibliography: Meta-science.bib
csl: mdpi.csl
abstract: ""
keywords: ""
output: 
  word_document:
    reference_docx: mdpi-formatting.docx
---

@@

  A scholarly communication system is supposed to serve five functions, but can do so in a narrow sense as it currently does, or in a wider sense. These functions of the scholarly communication system are (1) registration-, (2) certification-, (3) awareness-, and (4) archival [@roosendaal1997], and (5) incentives [@VandeSompel2004]. A narrow fulfillment of for example the registration function would mean that findings that are published are registered, but not all findings are registered (e.g., due to selective publication; [@10.1126/science.1255484]). Similarly, certification is supposed to occur through peer review, but peer review can exacerbate human biases in the assessment of quality (e.g., statistical significance increasing the perceived quality of methods; [@10.1007/bf01173636]).

  We propose an alternative design for scholarly communication based on modular research outputs with direct links between subsequent modules, forming a network. Whereas a paper-based approach communicates after a whole research cycle is completed, modular communication was proposed two decades ago [@Kircz1998;@VandeSompel2004;@10.7717/peerj-cs.78;@10.3233/ISU-2010-0613;@10.1163/095796511X560006;Nielsen2012]. These modules could be similar to sections of a research paper, but extend to modular research outputs such as software or materials. We propose to implement this modular communication on an "as-you-go" basis and include direct links to indicate provenance. This respects the chronological nature of research cycles and decreases the possibility for pernicious problems such as selective publication and making predictions after results are known (HARKing) [@Kerr1998].

@@

  Figure 1 shows a simple example of how these different modular research outputs (i.e., modules) would directly connect to each other. The connection between these modules only shows the direct descendance and could still include citations to other pieces of information. For example, a discussion module could be a direct descendant of a results module and could still include citations to other relevant findings. When one research cycle ends, a new one can link to the last module, continuing the chain of descendance. Incorporating the direct descendancy of these knowledge modules builds a different kind of network than citation and authorship networks. As such, this network would be an addition to these already existing citation and authorship networks; it does not seek to replace them.

![**Figure 1. Two Directed Acyclic Graphs (DAGs) of connected research stages. The ordering is chronological (top-bottom) and therefore modules that are situated below one another cannot refer upwards. Panel A shows a less complex network of modules; Panel B shows a more extensive network of modules.**](Topological_Ordering.png)

@@

  With immutable append-only registers, the chronology and content integrity of these outputs can be ensured and preserved over time. This can occur efficiently and elegantly with the Dat protocol (without a blockchain; [perma.cc/GC8X-VQ4K](https://perma.cc/GC8X-VQ4K)). In short, the Dat protocol is a peer-to-peer protocol (i.e., decentralized and openly accessible) that provides non-adjustable timestamps to each change that occurs within a folder, which is given a permanent unique address on the peer-to-peer Web (36^64 addresses possible) [@Dat2017]. The full details, implications, and potential implementations of this protocol for scholarly communication fall outside of the scope of this paper (see [https://dat-com-chris.hashbase.io](https://web.archive.org/web/20180316145316/https://dat-com-chris.hashbase.io/)).

>>>>

---
title: "As-you-go instead of after-the-fact: A network approach to scholarly communication and evaluation"
bibliography: Meta-science.bib
csl: mdpi.csl
abstract: ""
keywords: ""
output:
word_document:
reference_docx: mdpi-formatting.docx
---

@@

  A scholarly communication system is supposed to serve five functions, but can do so in a narrow sense as it currently does, or in a wider sense. These functions of the scholarly communication system are (1) registration-, (2) certification-, (3) awareness-, and (4) archival [@roosendaal1997], and (5) incentives [@10.1045/september2004-vandesompel]. A narrow fulfillment of for example the registration function would mean that findings that are published are registered, but not all findings are registered (e.g., due to selective publication; [@10.1126/science.1255484]). Similarly, certification is supposed to occur through peer review, but peer review can exacerbate human biases in the assessment of quality (e.g., statistical significance increasing the perceived quality of methods; [@10.1007/bf01173636]).

  We propose an alternative design for scholarly communication based on modular research outputs with direct links between subsequent modules, forming a network. Whereas a paper-based approach communicates after a whole research cycle is completed, modular communication was proposed two decades ago [@Kircz1998;@10.1045/september2004-vandesompel;@10.7717/peerj-cs.78;@10.3233/ISU-2010-0613;@10.1163/095796511X560006;@Nielsen2012]. These modules could be similar to sections of a research paper, but extend to modular research outputs such as software or materials. We propose to implement this modular communication on an "as-you-go" basis and include direct links to indicate provenance. This respects the chronological nature of research cycles and decreases the possibility for pernicious problems such as selective publication and making predictions after results are known (HARKing) [@Kerr1998].

@@

  Figure 1 shows two simple examples of how these different modular research outputs (i.e., modules) would directly connect to each other. The connection between these modules only shows the direct descendance and could still include citations to other pieces of information. For example, a discussion module could be a direct descendant of a results module and could still include citations to other relevant findings. When one research cycle ends, a new one can link to the last module, continuing the chain of descendance. Incorporating the direct descendancy of these knowledge modules builds a different kind of network than citation and authorship networks. As such, this network would be an addition to these already existing citation and authorship networks; it does not seek to replace them.

![**Figure 1. Two Directed Acyclic Graphs (DAGs) of connected research stages. The ordering is chronological (top-bottom) and therefore modules that are situated below one another cannot refer upwards. Panel A shows a less complex network of modules; Panel B shows a more extensive network of modules.**](Panel with DAGs.png)

@@

  With immutable append-only registers, the chronology and content integrity of these outputs can be ensured and preserved over time. This can occur efficiently and elegantly with the Dat protocol (without a blockchain; [perma.cc/GC8X-VQ4K](https://perma.cc/GC8X-VQ4K)). In short, the Dat protocol is a peer-to-peer protocol (i.e., decentralized and openly accessible) that provides non-adjustable timestamps to each change that occurs within a folder, which is given a permanent unique address on the peer-to-peer Web (36^64 addresses possible) [@Dat2017]. The full details, implications, and potential implementations of this protocol for scholarly communication fall outside of the scope of this paper (an extended technical explanation of the application of the Dat protocol can be found here [https://dat-com-chris.hashbase.io](https://web.archive.org/web/20180316145316/https://dat-com-chris.hashbase.io/)).
